### 2022-12-02 Documentation/Relatorio_2/Conclusion/Conclusion.tex
new project organization

\section{Conclusion}

For further steps the following tasks will be done:

- make experiments with different time windows. This will help to see the time needed for the system to stabilize.

- change power supply and oscilloscope libraries to an updated version wich covers both hardwares.

- correct and upgrade plots to make a better sense of the data.

- Include DHT11 temperature and humidity sensor to the setup.

- Thread sincronization in the main routine.