### 2023-01-17 Documentation/UFMG/modelo_PFC_latex/Resumo/Abstract.tex

\addcontentsline{toc}{chapter}{Abstract}

\begin{center}
\huge{{\bf Abstract}}
\vspace{2cm}
\end{center}

Write a version of your ``resumo'' in English. Beware of literal translations and double check the translation of technical terms.

\clearpage
\thispagestyle{empty}
\cleardoublepage