### 2023-02-06 Documentation/UFMG/Relatorio_4/Conclusion/Conclusion.tex
starting report 4

\section{Conclusion}

The integration with the pump machine in the automatic system brings a new dimension of possibilities in this project.
The goal at the moment is to improve the algorithm in order to be capable of scanning the stable region of cone jet mode in the Voltage X Flowrate map.
After that define the middle point in the stability island to set as a setpoint to our control system.
The control system will be capable of rejecting noises and disturbance in the system in order to keep always in the stable cone jet spraying mode.

For further goals we have:

 - Save the experiment data in a file while running the experiment. This will prevent lost of data if the program crashes and also not overflowing the program memory.

 - Install and run the experiments in a raspberry pi portable computer running Linux.

 - Study about the fourier transform peaks and how it can be used in the classification and in the mapping.

 - Improve the classification and routine performance.

 - Optimize the controller.

Until the moment the project has been shown good results.
The map routine, the control routine and the classification are already implemented and working.
Now its more about improvements to get better accuracy in results.