### 2023-02-06 Documentation/UFMG/Relatorio_4/Development/Development.tex

\section{Metodology and Results}

    The following subsections are about the work done in this 3rd part of the project.

\subsection*{2.1 Save experiment in real time}

\subsection*{2.1 Raspberry pi}

\subsection*{2.1 Correctly Join threads and qeues}