### 2023-03-03 Documentation/WT group/NHL/Capa/capa.tex
documentação

\begin{titlepage}
    \begin{center}

    \begin{figure}
            \centering
            % \subfigure{\includegraphics[width=5cm]{Capa/Figuras/wac.png}}
            \hspace{0.1\textwidth}
            \subfigure{\includegraphics[width=6cm]{Capa/Figuras/nhl-stenden.png}}
            \vspace{3cm} 
        \end{figure}

    {\bf\Large EHDA closed loop control system based on real time non-visual spray mode classification\\}
    \vspace{1cm} 
    {\Large Report}
    \vspace{2cm}

    %\hspace{0.3\textwidth} 
    {\bf\large Student: João Pedro Miranda Marques}\\
    \vspace{2cm}
    {\large Supervisor: Luewton L F Agostinho}
    \vspace{2cm}

    \today
    \vspace{2cm}

    %\hspace{0.3\textwidth} 
    \large \date{\today}
    \end{center}

    \end{titlepage}

\title{
    EHDA closed loop control system based on real time non-visual spray mode classification \\
    \large Relatório de Atividades}