### 2023-04-12 Documentation/UFMG/PFCII/Capa/capa.tex

\begin{titlepage}
    \begin{center}

    \vspace{2cm}
    \begin{figure}[h]
        \centering
        \includegraphics[scale=0.5]{Figuras/brasao_ufmg.png}
    \end{figure}
    \vspace{2cm}

    {\bf\Large EHDA closed loop control system based on real time non-visual spray mode classification\\}
    \vspace{2cm} 
    {\Large Monography}
    \vspace{2cm}

    %\hspace{0.3\textwidth} 
    {\bf\large Student: João Pedro Miranda Marques}\\
    \vspace{2cm}
    {\large Supervisors:\\ Luewton L F Agostinho\\
            Klaus Glanzer \\
            Antônio C. Munoz\\
            Vitor Angelo}
    \vspace{2cm}

    \today
    \vspace{2cm}

    %\hspace{0.3\textwidth} 
    \large \date{\today}
    \end{center}

    \end{titlepage}

    \newpage
\clearpage
\thispagestyle{empty}

\begin{titlepage}

\centering
\vspace{2cm}
\textbf{Monography}\\
\vspace{2cm}
\centering
\textbf{EHDA closed loop control system based on real time non-visual spray mode classification.}\\
\vspace{5cm}

\parbox{1.0\textwidth} 
{\large 
Document submitted to the examining board designated by the Lectoraat Watertechnology at the NHL Stenden University of Applied Sciences and the Colegiado de Engenharia de Controle e Automação at the Universidade Federal de Minas Gerais as part of the requirements for approval in the subject Final Course Project II.}

\vspace{7cm}

\end{titlepage}

\title{
    EHDA closed loop control system based on real time non-visual spray mode classification \\
    \large Monography}