### 2023-04-12 Documentation/UFMG/PFCII/acknowledgment/acknowledgment.tex

\addcontentsline{toc}{chapter}{Acknowledgment}

\begin{center}
\huge{{\bf Acknowledgments}}
\vspace{4cm}
\end{center}

I wish a big thanks to my mother who always supported me and made me believe in my potential. Teaching me dedication, patience and discipline.
To my father who taught me not to be satisfied and always looking for more, that's what I call ambition and energy to always try to get out of the comfort zone.
To my dear friends from República Casassanta for making university life such a unique and unforgettable moment.

\clearpage
\thispagestyle{empty}
\cleardoublepage