### 2023-04-22 Documentation/UFMG/PFCII/Conclusion/Conclusion.tex
documentation

\chapter{Conclusion}

In conclusion, the goals of this project have been largely accomplished. The implementation of a Python routine for complete automation and integration of the system has significantly improved the precision of experiments and enabled real-time collection of extensive data without requiring the expertise and time of a specialized scientist. The success of this project underscores the potential of automation in enhancing scientific research and accelerating progress towards achieving research goals.
In addition to its potential in scientific research, the software developed in this project can also be leveraged by industrial approaches. Furthermore, the classification methods and system control implemented in this project have yielded promising results and can be further optimized in the future. Overall, this project serves as a valuable contribution towards the advancement of automated systems and the optimization of experimental processes, with broad implications for both research and industrial applications.

\section{Proposal for continuing}

\subsection{Fuzzy Controller}

        The fuzzy approach of controller was explored and simulated but not used in the final version of the project.
        This because for this fuzzy approach we need to have the input variables for the fuzzy machine to be fuzzyfied.
        Which means that to use a fuzzy logic in our controller loop the classification must be fuzzyfied and our classification algorithm was not developed in order to give a classification and its current membership function.

        For that, I tried to fuzzyfy the controller input by the data acquired in the step routine. With the data I mapped the area of each spraying mode according to its potential and fuzzyfied this as shown in the Figure X.

        \begin{multicols}{2}

            \begin{figure}[H]
                \centering
                \resizebox{90mm}{!}{\includegraphics{Figuras/fuzzy/fuzzyfy_input.png}}
                \caption{Fuzzyfication}
                \label{fig:fuzzy_input}
            \end{figure}

            \begin{figure}[H]
                \centering
                \resizebox{90mm}{!}{\includegraphics{Figuras/fuzzy/defuzzyfy.png}}
                \caption{Defuzzyfication}
                \label{fig:fuzzy_output}
            \end{figure}

        \end{multicols}

        \begin{figure}[H]
            \centering
            \resizebox{80mm}{!}{\includegraphics{Figuras/fuzzy/rules.png}}
            \caption{Fuzzy Rules}
            \label{fig:fuzzy_rules}
        \end{figure}

        \begin{multicols}{2}

            \begin{figure}[H]
                \centering
                \resizebox{90mm}{!}{\includegraphics{Figuras/fuzzy/test1.png}}
                \caption{Test 1: fuzzy controller}
                \label{fig:fuzzy_test1}
            \end{figure}

            \begin{figure}[H]
                \centering
                \resizebox{90mm}{!}{\includegraphics{Figuras/fuzzy/test2.png}}
                \caption{Test 2: fuzzy controller}
                \label{fig:fuzzy_test2}
            \end{figure}

        \end{multicols}

\clearpage