### 2023-04-22 Documentation/UFMG/PFCII/Apendices/ApendiceA.tex
documentation

\chapter{Trials and failures}

\section{Machine Learn}

I've tried to create a database with the samples of current. Each sample was a vector with 50k values corresponding a 0.5s of data collected in a 100KHz oscilloscope. And in a second column its

\subsection{Spectogram analysis}