### 2023-04-22 Documentation/UFMG/PFCII/Apendices/ApendiceB.tex

\chapter{O que mais faltou}

Inclua aqui informações que não sejam tão relevantes para o entendimento do projeto mas que ainda sejam importantes para documentá-lo.

put here the stability difficult about doing experiments

- syringe pump inclined
- pumping liquid to avoid bubbles
- mechanical noise in the pumping machine
- external electrical noise collected by antennas or connections

Foto e detalhes dos instrumentos

\begin{figure}[H]
    \centering
    \resizebox{150mm}{!}{\includegraphics{Figuras/naked_eyes.png}}
    \caption{EHDA physical concept }
\end{figure}