### 2023-04-22 Documentation/UFMG/PFCII/Apendices/ApendiceA.tex
documentation

\chapter{Trials and failures}

\section{Machine Learning}

I've tried to create a database with the samples of current. Each sample was a vector with 50k values corresponding a 0.5s of data collected in a 100KHz oscilloscope. And in a second column its

\subsection{Spectogram analysis}

\begin{figure}[H]
    \center
    \includegraphics[width=3cm]{Figuras/spectogram/5_Dripping.png}
    \label{fig:spectrogram1}
    \caption{Dripping}
  \end{figure}

%   \begin{figure}[H]
%     \center
%     \includegraphics[width=3cm]{Figuras/spectogram/5_Dripping.png}
%     \label{fig:spectrogram1}
%     \caption{Dripping}
%   \end{figure}